opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, cache=FALSE)

---
name: easier-models
layout: post
title: Easier model diagnostics function
date: 2012-12-10
author: Scott Chamberlain
tags: 
- R
- data
- model
- statistics
- ggplot2
---

It's a pain to write a bunch of lines of code to do model diagnostics. This little function does a nice job of putting out a number of diagnostic plots, and the gives the ANOVA table too, with an option to just write the ANOVA table out as a data.frame so you can easily combine many ANOVA tables for easy writing to CSV for instance. 

***************

### The function
```{r thefunc }
# install.packages(c("car","ggplot2"))
library(car); library(ggplot2)

fitmod <- function(equation, data, gimme=FALSE) {
	conn_mod1 <- lm(equation, data=data) # run model
	m1 <- fortify(conn_mod1, data) # adds model residuals/etc to original data.frame
	anov <- Anova(conn_mod1, type=3) # so we can get type III SS
	a <- ggplot(m1, aes_string(x = as.character(equation)[[2]]) ) + geom_histogram()
	b <- qplot(.fitted, .stdresid, data=m1, geom="point")
	c <- qplot(.stdresid, data=m1, geom="histogram")
	if(gimme){
		temp <- data.frame(anov[1:4])
		temp$vars <- row.names(temp)
		temp
	} else
	{
		list(list(anov),
				 do.call(grid.arrange,  list(a,b,c))
		)
	}
}
````

***************

### Just run the model and give results and diagnostic plots. This gives a histogram of the response variable, a plot of the fitted numbers vs. residuals, and a histogram of the residuals.
```{r outputall }
fitmod(Sepal.Length ~ Sepal.Width * Species, iris)
````

***************

### Or you can run the model and get just the ANOVA table back as a data.frame that can be combined with others, e.g., to write to a CSV file
```{r gimme }
library(plyr)
# Simple example
fitmod(Sepal.Length ~ Sepal.Width * Species, iris, gimme=TRUE)

# Or do many in lapply call to combine them
mymods <- list(Sepal.Length ~ Sepal.Width * Species, Petal.Length ~ Petal.Width * Species)
out <- llply(mymods, function(x) fitmod(x, iris, TRUE))
names(out) <- c("mod1","mod2")
ldply(out)
```


*********
#### Get the .Rmd file used to create this post [at my github account](https://github.com/sckott/scott/blob/gh-pages/_drafts/2012-12-10-easier-models.Rmd) - or [.md file](https://github.com/sckott/scott/blob/gh-pages/_posts/2012-12-10-easier-models.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).